Binning Strategies and Related Loss for Large Data

### 2015-05-06 Paper.Rnw
striping

@@

If we view the binned scatterplot as a visual estimator of the bivariate density then we may consider a few desirable properties of estimators: unbiasedness, consistency and efficiency. \ktm{Binned scatterplots are inherently bias displays of spatial information as they shift visual emphasis from true location of individual points to the geometric centers of tiles. However, as bin size becomes increasingly fine, the plot will place visual emphasis more precise spatial locations. When minimal binning occurs -- with one unique coordinate pair at the center of each bin -- the density estimate then spatially exact; thus making the binned scatterplot a consistent visual estimator. We may also consider minimally binning as a spatially efficient estimator because it minimizes spatial loss in the visual density estimator. It is worth noting that a density estimate requires the combination of spatial and frequency information and that the estimation properties were considered only through the lens of spatial loss. This makes the assumption that frequency information is rendered through a precise continuous mapping of bin frequencies to bin shades. %Based on the conflicting effects on the estimation properties, of bin size can be seen as a smoother for density estimate that can be adjusted to cut through the noise in the data, but as with all data smoothers caution should be used so that fine signal is not destroyed.  
}

>>>>

\hh{OK, better - I still don't understand the 'but' at the top of the paragraph. There seems to be some implicit negation, which I can't see it. Could you make this explicit?}

@@

If we view the binned scatterplot as a visual estimator of the bivariate density then we may consider a few desirable properties of estimators: unbiasedness, consistency and efficiency. Binned scatterplots are inherently biased displays of spatial information as they shift visual emphasis from the true location of individual points to the geometric centers of tiles. However, as bin sizes become increasingly fine,  the bias decreases as the plot shows more precise spatial locations.
\ktm{When minimal binning occurs -- with one unique coordinate pair at the center of each bin -- the density estimate then spatially exact; thus making the binned scatterplot a consistent visual estimator. We may also consider minimally binning as a spatially efficient estimator because it minimizes spatial loss in the visual density estimator. It is worth noting that a density estimate requires the combination of spatial and frequency information and that the estimation properties were considered only through the lens of spatial loss. This makes the assumption that frequency information is rendered through a precise continuous mapping of bin frequencies to bin shades. %Based on the conflicting effects on the estimation properties, of bin size can be seen as a smoother for density estimate that can be adjusted to cut through the noise in the data, but as with all data smoothers caution should be used so that fine signal is not destroyed.  
}
